---
title: "HARP Data Quality Analysis"
author: "Victor Verma"
date: 2024-08-02
date-format: iso
published-title: "Date Created"
format:
  html:
    embed-resources: true
    toc: true
    toc-location: left
    number_sections: true
    theme: united
---

```{r}
#| label: setup
#| include: false

here::i_am("notebooks/quality_analysis/quality_analysis.qmd")

library(arrow)
library(cowplot)
library(gridExtra)
library(here)
library(kableExtra)
library(knitr)
library(tidyverse)
library(vctrs)

harp_tbl <- read_parquet(here("data/processed/harp_data.parquet"))
```

In this analysis, we investigate quality issues in the HARP data for the period `r str_c(format(range(harp_tbl$T_REC), "%F"), collapse = " to ")`.

# Use Only Solar Maximum Data

For times between the first and last times in the data, we flag those that lie within the lifetime of some HARP.
```{r}
time_tbl <- harp_tbl %>%
  distinct(T_REC) %>%
  add_column(is_in_harp_lifetime = TRUE) %>%
  complete(T_REC = full_seq(T_REC, period = 720)) %>%
  mutate(is_in_harp_lifetime = coalesce(is_in_harp_lifetime, FALSE))
```

The majority of times are within some HARP lifetime, but a significant minority aren't.
```{r}
time_tbl %>%
  count(is_in_harp_lifetime, sort = TRUE) %>%
  mutate(prop = n / sum(n)) %>%
  kable(
    "html",
    digits = 2,
    col.names = c(
      "Is in a HARP Lifetime?", "Number of Times", "Proportion of Times"
    ),
    align = "crr",
    format.args = list(big.mark = ",")
  ) %>%
  kable_styling()
```

The sequence of times can be partitioned into runs such that within each run, either every time is in some HARP lifetime or no time is in some HARP lifetime. For each run, we compute its start and end times and its duration.
```{r}
#| message: false

time_rle_obj <- rle(time_tbl$is_in_harp_lifetime)
time_summary_tbl <- time_tbl %>%
  add_column(
    run_id = rep(seq_along(time_rle_obj$values), times = time_rle_obj$lengths)
  ) %>%
  group_by(run_id, is_in_harp_lifetime) %>%
  summarize(
    start_time = min(T_REC),
    end_time = max(T_REC),
    days_elapsed = difftime(max(T_REC), min(T_REC), units = "days"),
    num_times = length(T_REC)
  ) %>%
  ungroup() %>%
  select(!run_id)
```

In the plot below, each dot represents one run of times. A run is blue if it consists exclusively of times that lie within the lifetime of some HARP; a run is orange if it only contains times that do not have that property. The first coordinate of a dot is the start time of the corresponding run, while the second coordinate is the run's duration in days. Orange runs should not be used as there is no HARP data for them and no way to impute it. It seems reasonable to train a model with at least a few months' worth of data; there are only two blue runs that could be divided into training sets with that much data. Both of those runs correspond to solar maxima.
```{r}
#| label: run_length_vs_start_time

time_summary_tbl %>%
  mutate(days_elapsed = as.double(days_elapsed)) %>%
  ggplot(aes(start_time, days_elapsed, color = is_in_harp_lifetime)) +
  geom_point() +
  scale_y_log10() +
  scale_color_manual(
    labels = c(
      `TRUE` = "Every Time Is in a HARP Lifetime",
      `FALSE` = "No Time Is in a HARP Lifetime"
    ),
    values = c(`TRUE` = "blue", `FALSE` = "orange")
  ) +
  labs(x = "Run Start Time", y = "Run Length (Days)", color = "") +
  theme_bw() +
  theme(legend.position = "top")
```

The table below displays the five longest runs that have HARP data at each time. As stated above, only two runs are long enough that they can be carved into multiple reasonably-sized training sets.
```{r}
time_summary_tbl %>%
  filter(is_in_harp_lifetime) %>%
  select(!is_in_harp_lifetime) %>%
  slice_max(days_elapsed, n = 5) %>%
  mutate(days_elapsed = as.double(days_elapsed)) %>%
  kable(
    "html",
    digits = 2,
    col.names = c(
      "Run Start Time", "Run End Time",
      "Run Length (Days)", "Run Length (Times)"
    ),
    align = "ccrr",
    format.args = list(big.mark = ",")
  ) %>%
  kable_styling()
```

In the sequel, we will only consider times from the longest run.
```{r}
longest_run <- time_summary_tbl %>%
  filter(is_in_harp_lifetime) %>%
  slice_max(days_elapsed)
harp_tbl <- filter(
  harp_tbl, between(T_REC, longest_run$start_time, longest_run$end_time)
)
```

# Problems in the Solar Maximum Data

The vector below will prove to be useful. The list of SHARP parameters was obtained from [here](http://jsoc.stanford.edu/doc/data/hmi/sharp/sharp.htm).
```{r}
sharp_params <- c(
  "USFLUX", "MEANGAM", "MEANGBT", "MEANGBZ", "MEANGBH", "MEANJZD", "TOTUSJZ",
  "MEANALP", "MEANJZH", "TOTUSJH", "ABSNJZH", "SAVNCPP", "MEANPOT", "TOTPOT",
  "MEANSHR", "SHRGT45"
)
```

A HARP record can have one or more problems. A record can be missing a value for a SHARP parameter; we call such records incomplete. If a record is missing values for all SHARP parameters, we call it missing. For certain records, some SHARP parameter is infinite. Finally, some records are marked as low-quality. We flag records with these problems. We also flag records for which part or all of the bounding box was more than 68 degrees from the central meridian. Image quality degrades near the limb of the Sun, so the problems for a record could be due to the bounding box being far from the central meridian.
```{r}
na_flags <- harp_tbl %>%
  select(all_of(sharp_params)) %>%
  transmute(
    is_rec_incomplete = !vec_detect_complete(.) & !vec_detect_missing(.),
    is_rec_missing = vec_detect_missing(.)
  )
inf_flags <- harp_tbl %>%
  select(all_of(sharp_params)) %>%
  apply(1, function(r) any(is.infinite(r), na.rm = TRUE))
harp_tbl <- harp_tbl %>%
  bind_cols(na_flags) %>%
  add_column(has_inf = inf_flags) %>%
  mutate(
    is_low_quality = QUALITY != "0x00000000",
    has_low_lon = LON_MIN < -68, has_high_lon = LON_MAX > 68
  )
```

Most records don't have any of the problems described above, but a sizable fraction have some problem. Only `r signif(100 * mean(with(harp_tbl, is_rec_missing + is_rec_incomplete + has_inf + coalesce(is_low_quality, FALSE)) > 1), 2)`\% of records have more than one problem. The most important subsets of problematic records consist of records that are only missing (`r signif(100 * mean(harp_tbl$is_rec_missing), 2)`\% of records), records that are only low-quality (`r signif(100 * mean(coalesce(harp_tbl$is_low_quality, FALSE)), 2)`\% of records), and records that are only incomplete (`r signif(100 * mean(harp_tbl$is_rec_incomplete), 2)`\% of records). Records with an infinite value are virtually nonexistent.
```{r}
harp_tbl %>%
  count(
    is_rec_missing, is_rec_incomplete, has_inf, is_low_quality, sort = TRUE
  ) %>%
  mutate(prop = n / sum(n)) %>%
  kable(
    "html",
    digits = 2,
    col.names = c(
      "Is Record Missing?", "Is Record Incomplete?",
      "Does Record Have an \u221e?",
      "Is Record Low-Quality?",
      "Number of Records", "Proportion of Records"
    ),
    align = "cccrr",
    format.args = list(big.mark = ",")
  ) %>%
  kable_styling()
```

# A Possible Cause of the Problems: HARP Age

It could be that when a HARP is near the beginning or close to the end of its lifetime, it's difficult to detect, causing data quality problems. We investigate this possibility in this section. For each HARP, for each time in its lifespan, we can compute the corresponding lifespan fraction, with the first time corresponding to a lifespan fraction of zero, and the last time corresponding to a lifespan fraction of one.
```{r}
harp_tbl <- harp_tbl %>%
  group_by(HARPNUM) %>%
  mutate(lifespan_frac = seq(0, 1, length.out = n())) %>%
  ungroup()
```

Next, across all HARPs, we can bin the various lifespan fraction values, and for each bin, we can compute the proportions of corresponding records that are missing, incomplete, and low-quality. This is done below.
```{r}
#| message: false

breaks <- seq(0, 1, by = 0.05)
midpoints <- (breaks[-length(breaks)] + breaks[-1]) / 2
breaks[length(breaks)] <- Inf
prop_tbl <- harp_tbl %>%
  mutate(
    lifespan_frac = cut(lifespan_frac, breaks = breaks, right = FALSE),
    midpoint = midpoints[as.integer(lifespan_frac)]
  ) %>%
  group_by(lifespan_frac, midpoint) %>%
  summarize(
    missing_rec_prop = mean(is_rec_missing),
    incomplete_rec_prop = mean(is_rec_incomplete),
    low_quality_prop = mean(is_low_quality, na.rm = TRUE)
  ) %>%
  ungroup()
```

If youth or old age causes data quality problems, the various proportions computed in the previous chunk should decrease and then increase as the lifespan fraction increases. Below, we plot the proportions versus the midpoints of the lifespan fraction bins. We see exactly the pattern we would expect in the incomplete record curve. Youth doesn't appear to contribute to missingness, but old age appears to; the proportion of missing records rises as the lifespan fraction increases, with the rate of growth higher for larger lifespan fractions. Age has no effect on the low-quality proportion.
```{r}
#| label: prop_plot

prop_tbl %>%
  pivot_longer(
    cols = !c(lifespan_frac, midpoint),
    names_to = "prop_type", names_pattern = "([a-z_]+)_prop",
    values_to = "prop_val"
  ) %>%
  ggplot(aes(midpoint, prop_val, linetype = prop_type)) +
  geom_line() +
  scale_linetype_manual(
    labels = c(
      missing_rec = "Missing Record",
      incomplete_rec = "Incomplete Record",
      low_quality = "Low-Quality Record"
    ),
    values = c("solid", "dashed", "dotted")
  ) +
  labs(
    x = "Lifespan Fraction", y = "Proportion Value",
    linetype = "Proportion Type"
  ) +
  theme_bw() +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.15, 0.82),
    legend.background = element_rect(fill = NA)
  )
```

# Investigations into the Problems

It's most common for an incomplete record to be missing a value for ten SHARP parameters.
```{r}
harp_tbl %>%
  filter(is_rec_incomplete) %>%
  select(all_of(sharp_params)) %>%
  add_column(num_nas = apply(., 1, function(r) sum(is.na(r)))) %>%
  count(num_nas, sort = TRUE) %>%
  mutate(prop = n / sum(n)) %>%
  kable(
    "html",
    digits = 2,
    col.names = c("Number of NAs", "Count", "Proportion"),
    align = "rrr",
    format.args = list(big.mark = ",")
  ) %>%
  kable_styling()
```

All of the incomplete records that are missing values for ten SHARP parameters look the same. For these records, the SHARP parameters that do have values equal zero.
```{r}
harp_tbl %>%
  filter(is_rec_incomplete) %>%
  select(all_of(sharp_params)) %>%
  add_column(num_nas = apply(., 1, function(r) sum(is.na(r)))) %>%
  count(across(c(all_of(sharp_params), num_nas)), sort = TRUE) %>%
  slice_head(n = 1) %>%
  kable(
    "html",
    col.names = c(sharp_params, "Number of NAs", "Count"),
    align = rep("r", 18),
    format.args = list(big.mark = ",")
  ) %>%
  kable_styling()
```

```{r}
knitr::opts_chunk$set(eval = FALSE)
```

The SHARP parameters that do have values for the records considered in the previous chunk are `ABSNJZH`, `SAVNCPP`, `TOTPOT`, `TOTUSJH`, `TOTUSJZ`, and `USFLUX`. The histograms below, which are based on complete, high-quality records from close to the central meridian, suggest that zero is an invalid value for these SHARP parameters, though the evidence of this for `ABSNJZH` is relatively weak. However, it isn't clear whether this is true from the descriptions at [this link](http://jsoc.stanford.edu/doc/data/hmi/sharp/sharp.htm).
```{r}
#| message: FALSE
#| warning: FALSE

set.seed(1)
harp_tbl %>%
  filter(
    is_rec_complete,
    abs(LON_MAX) <= 68, abs(LON_MIN) <= 68,
    QUALITY == "0x00000000"
  ) %>%
  select(all_of(sharp_params)) %>%
  mutate(across(everything(), abs)) %>%
  pivot_longer(
    everything(), names_to = "sharp_param", values_to = "abs_val"
  ) %>%
  group_by(sharp_param) %>%
  slice_sample(n = 100000) %>%
  ungroup() %>%
  ggplot(aes(abs_val)) +
  facet_wrap(vars(sharp_param), scales = "free_x") +
  geom_histogram() +
  scale_x_log10() +
  labs(x = "Absolute Value", y = "Number of Records") +
  theme_bw() +
  theme(axis.text.x = element_text(vjust = 0.3, angle = 30))
```

Among the incomplete records that are missing just one value, the missing values are concentrated in just two SHARP parameters. Most of these records are missing a value for `MEANSHR`, but some are missing a value for `MEANGAM`.
```{r}
harp_tbl %>%
  filter(is_rec_incomplete) %>%
  select(all_of(sharp_params)) %>%
  add_column(num_nas = apply(., 1, function(r) sum(is.na(r)))) %>%
  filter(num_nas == 1) %>%
  select(!num_nas) %>%
  summarize(across(everything(), ~ mean(is.na(.x)))) %>%
  kable("html", digits = 2) %>%
  kable_styling()
```

Half of the incomplete records are missing their maximum and minimum longitudes;
the other half have both.
```{r}
harp_tbl %>%
  filter(is_rec_incomplete) %>%
  mutate(is_lon_max_na = is.na(LON_MAX), is_lon_min_na = is.na(LON_MIN)) %>%
  count(is_lon_max_na, is_lon_min_na, sort = TRUE) %>%
  mutate(prop = n / sum(n)) %>%
  kable(
    "html",
    digits = 2,
    col.names = c(
      "Is Maximum Longitude Missing?", "Is Minimum Longitude Missing?",
      "Count", "Proportion"
    ),
    format.args = list(big.mark = ",")
  ) %>%
  kable_styling()
```

Among the incomplete records that have both longitudes, there don't appear to be characteristic values for the longitudes, though there are concentrations of records with small values for both or large values for both.
```{r}
harp_tbl %>%
  filter(is_rec_incomplete, !is.na(LON_MAX), !is.na(LON_MIN)) %>%
  ggplot(aes(LON_MIN, LON_MAX)) +
  geom_point(alpha = 0.3) +
  labs(x = "Minimum Longitude", y = "Maximum Longitude") +
  theme_bw()
```

Strangely, almost all the incomplete records have a `QUALITY` value of `0x00000000`, which is supposed to be the value for records that do not have problems.
```{r}
harp_tbl %>%
  filter(is_rec_incomplete) %>%
  count(QUALITY, sort = TRUE) %>%
  mutate(prop = n / sum(n)) %>%
  kable(
    "html",
    digits = 2,
    col.names = c("QUALITY", "Count", "Proportion"),
    format.args = list(big.mark = ",")
  ) %>%
  kable_styling()
```

For a given HARP, make a tibble with one row for each run of consecutive complete records and one row for each run of consecutive incomplete or missing records, i.e., records to which an imputation algorithm could be applied.
```{r}
make_harp_run_summary_tbl <- function(single_harp_tbl) {
  single_harp_tbl %>%
    pull(is_rec_complete) %>%
    rle() %>%
    unclass() %>%
    as_tibble() %>%
    rename(length = lengths, val_type = values) %>%
    mutate(val_type = if_else(val_type, "Complete", "Incomplete/Missing"))
}
```

Make a run summary tibble for each HARP.
```{r}
harp_run_summary_tbls <- harp_tbl %>%
  nest(.by = HARPNUM, .key = "single_harp_tbl") %>%
  mutate(
    num_times = map_int(single_harp_tbl, nrow),
    run_summary_tbl = map(single_harp_tbl, make_harp_run_summary_tbl),
    num_runs = map_int(run_summary_tbl, nrow),
    times_per_run = num_times / num_runs
  )
```

For the typical HARP, the number of runs is in the low dozens, and the average number of times or records per run is also in the low dozens. The strange sawtooth pattern in the left histogram is due to the fact that a HARP is more likely to have an odd number of runs than an even number. This is because the first and last records should typically be complete, and thus the first and last runs should be runs of complete records; if this is the case, then the number of runs is odd.
```{r}
#| message: FALSE

panel1 <- harp_run_summary_tbls %>%
  ggplot(aes(num_runs)) +
  geom_histogram(binwidth = 1, center = 0) +
  labs(x = "Number of Runs", y = "Number of HARPs") +
  theme_bw()
panel2 <- harp_run_summary_tbls %>%
  ggplot(aes(times_per_run)) +
  geom_histogram() +
  labs(x = "Average Number of Times per Run", y = "Number of HARPs") +
  theme_bw()
grid.arrange(panel1, panel2, nrow = 1)
```

**The main takeaway from the preceding analysis is that the missing values are concentrated in certain rows instead of being scattered across the dataset, which is what would be best for a multiple imputation algorithm like MICE. The rows the missing values lie in tend to occur in runs, making the imputation problem even more challenging.**
